---
title: Popular songs (Taylor's Version)
author: Yanely Luna
date: '2022-02-24'
slug: []
categories: []
tags: []
subtitle: ''
excerpt: 'Exploring songs by my favorite artist in the world.'
images: ~
series: ~
layout: single
---

If you have known me in person, I think it would take you less than ten minutes to realize that I love Taylor Swift's music (and work in general) more that any music from any other artist ever. In that case, is no surprise that I dedicated this post to analyse her discography data. I will be using a dataset from [kaggle](https://www.kaggle.com/thespacefreak/taylor-swift-spotify-data) to first do some exploration and then to try to determine the song's popularity based on its properties.

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE,message = FALSE)
colores <- c("#65D267","#A665D2","#7A55F7","#F77A55","#F75581","#CB55F7","#5581F7", #3:7 rating
             "#DCF755","#55DCF7","#F7555E","#F7EE55") 
```

# Libraries

```{r lib}
library(dplyr) # For manipulating data
library(tidyr) # same
library(ggplot2) # For graphs
library(gridExtra) # same
library(ggtextures) # For textured graphs
```

# Dataset

```{r dataset}
taylor <- read.csv("spotify_taylorswift.csv")
str(taylor)

```

This dataset contains `r nrow(taylor)` observations from 16 variables, which according to the source represent the following:

+ `name`: The name of the song
+ `album`: Name of the album
+ `artist`: Name of artist/s involved
+ `release_date`: Release date of album
+ `length`: Song length in milliseconds
+ `popularity`: Percent popularity of the song based on Spotify's algorithm (posibly the number of stream at a certain period of time).
+ `danceability`: How suitable a track is for dancing based on a combination of musical elements.
+ `acousticness`: How acoustic a song is.
+ `energy`: A perceptual measure of intensity and activity. 
+ `instrumentalness`: The amount of vocals in the song.
+ `liveness`: Probability that the song was recorded with a live audience.
+ `loudness`: Tendency of music to be recorded at steadily higher volumes.
+ `speechiness`: Presence of spoken words in a track (if the speechiness of a song is above 0.66, it is probably made of spoken words, a score between 0.33 and 0.66 is a song that may contain both music and words, and a score below 0.33 means the song does not have any speech)
+ `valence`: A measure of how sad or happy the song sounds.
+ `tempo`: Beats per minute.

We can drop the variable `X` since it does not give us information and look for other variables that we may not need.

When we take a closer look to the _folklore_ album (or any other album) we notice that the `artist` column has only one value even when other artist participate on the song (e.g. _exile_). We check that is the case for the whole dataset, so we can drop that variable later.

```{r summary}
taylor %>% count(album)

taylor %>% filter(album=="folklore (deluxe version)") %>% select(name, artist)

taylor %>% count(artist)
```

We know each of the nine albums in this dataset were released on different dates so we don't need the `release_date` variable for now.

```{r release_date}
taylor %>% group_by(album) %>% count(release_date)

#drop variables
taylor <- taylor %>% select(-X,-artist,-release_date)

# Change the album names to be shorter and rearrange them according to realse date.
taylor$album <- recode_factor(taylor$album, "Taylor Swift"="Debut",
                              "Fearless (Taylor's Version)" = "FearlessTV",
                              "Speak Now (Deluxe Package)" = "Speak Now",
                              "Red (Deluxe Edition)" = "Red",
                              "1989 (Deluxe)" = "1989",
                              "reputation"="reputation",
                              "Lover"="Lover",
                              "folklore (deluxe version)" = "folklore",
                              "evermore (deluxe version)"="evermore")
```

# Exploratory Data Analysis

Now, we can start exploring the numeric variables. Firstly, the `popularity` variable has a mean of `r round(mean(taylor$popularity),2)` and we notice some songs has a value of 0. It turns out, these are tracks included in the deluxe version of _1989_ but they are not actual songs. I chose to drop these three observations.

```{r popularity}
ggplot(taylor, aes(x=popularity)) +
  geom_histogram(fill=colores[2],binwidth = 3) +
  theme_classic() + ggtitle("Popularity distribution")

summary(taylor$popularity)

filter(taylor, popularity==0) %>% select(name, album, popularity)

taylor <- taylor %>% filter(popularity>0)

ggplot(taylor, aes(x=popularity)) +
  geom_histogram(fill=colores[2],binwidth = 3) +
  theme_classic() + ggtitle("Popularity distribution (without non-songs tracks)")
```

We can compare the popularity of each album by taking the mean of its songs' popularity (and let's add some glitter).

```{r geom_textured}
img = "https://t3.ftcdn.net/jpg/04/44/39/80/360_F_444398090_ek1YrGhZa2AZmiUZwALuMe3cRafivme9.jpg"
taylor %>% group_by(album) %>% 
  summarise(album_popularity=round(mean(popularity),2)) %>%
  ggplot(aes(x=album,y=album_popularity)) +
  geom_textured_col(image = img, color = "white", width = 0.8) +
  geom_text(aes(label=album_popularity),size=4, nudge_y = 5) +
  theme_classic() +
  ggtitle("Album's Popularity") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  coord_flip()
```

Now, let's have a quick look at the distribution of the other numeric variables.

```{r tay_long}
tay_long <- taylor %>%
  select(popularity,length, instrumentalness, danceability,acousticness,energy,liveness,loudness,
         speechiness,valence, tempo) %>% 
  pivot_longer(cols = length:tempo)

ggplot(tay_long, aes(x=value)) +
  geom_histogram(aes(fill=name)) +
  facet_wrap(~name, scales = "free") +
  theme_classic() + 
  theme(legend.position = "none") +
  scale_fill_manual(values = colores)
```

Let's take a closer look at `instrumentalness` since it seems most of its values are zero.

```{r instr}
summary(taylor$acousticness)

ggplot(taylor, aes(instrumentalness, popularity)) +
  geom_point(color=colores[5]) +
  theme_classic()
```
We can see that the values of `instrumentalness` are very close to zero and the variable doesn't seem to influence our response (`popularity`) so we decide to drop it.

Now we explore the correlation between the remaining continuous variables. We notice that `energy` and `loudness` have the strongest positive correlation of all pairs of variables; meanwhile, `loudness` and `acousticness` have the strongest negative correlation. This could be an indicative of collinearity problems for a regression model. 

```{r cor}
library(corrplot)
cont_vars <- c(3:7,9:13)
taylor_cont_vars <- taylor[,cont_vars]

cor_m <- cor(taylor_cont_vars)
corrplot(cor_m, method = 'shade', order = 'AOE', type = "lower")
```

Let's look at the other continuous variables compared with `popularity`. These scatter plots give us an idea of how much does a certain variable influences the response variable. We notice that `loudness` and `length` seem to have a negative relation with `popularity`; meanwhile, `speechiness` and `danceability` seem to have a positive one.

```{r pairs}
tay_long <- filter(tay_long, name!= "instrumentalness") #%>% 
  #mutate(value = ifelse(name=="length",format(value, scientific = TRUE), value))

ggplot(tay_long, aes(x=value, y=popularity)) +
  geom_point(aes(color=name)) +
  geom_smooth(method="lm", se=FALSE) +
  facet_wrap(~name, scales = "free") +
  theme_classic() + 
  theme(legend.position = "none",
        axis.text.x = element_text(size=7)) +
  scale_fill_manual(values = colores)

```

# Modeling

We can start by fitting a multiple linear regression model with all continuos variables.

```{r rlm}
mod_rlm1 <- lm(popularity~., data = taylor_cont_vars)
summary(mod_rlm1)

step_mod1 <- step(mod_rlm1, direction = "both")

mod2 <- lm(popularity ~ length + acousticness + loudness + 
    speechiness, data = taylor_cont_vars)

summary(mod2)

```

```{r}
mod_rlm2 <- lm(popularity~., data = select(taylor,-name,-instrumentalness))
summary(mod_rlm2)

step(mod_rlm2, direction = "both")

mod4 <- lm(popularity ~ album + liveness + loudness + valence + tempo, data = taylor)

summary(mod4)

```

